---
title: "mutType (R Project)"
subtitle: |
  | Determine the mutation type for a set of single nucleotide variants in a genome
  |
  | Alessandro Giulivo
  | Scientific Programming - Prof Rosario M. Piro
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mutType_vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE
)
```

```{r setup}
library(mutType)
```

## Description

The mutType package takes a set of mutations (single nucleotide variants, SNVs)
in VCF format,
and determines for each mutation the corresponding mutation type.  
Check out the [full github repository](https://github.com/AlessandroGiulivo/sciprog/tree/main/RProj).


## mutType function

The mutType function compares a list of mutations in VCF format to a reference
genome and returns the mutation type along with upstream and downstream
bases for a total length = `context_length`.

**Usage:** `mutType(VCFFile, refGenome, context_length, graphics = TRUE)`

- `VCFFile`: a set of mutations in VCF Format
- `refGenome`: Full reference genome sequences as provided by
UCSC in Biostrings objects
- `context_length`: an odd integer representing the length of region
around the mutation
- `graphics`: default = TRUE; if TRUE, saves a barplot of mutation type
frequencies in a pdf file
**returns** a data.frame object containing mutations from `VCFFile` along with
mutation types as "UP[REF>ALT]DOWN"

**Example**
```{r mutType_example}
library(BSgenome.Hsapiens.UCSC.hg19)
Hs <- Hsapiens

sample <- system.file("extdata", "sample.vcf", package = "mutType")

muts <- mutType(sample, Hs, 7, graphics = FALSE)

muts
```


## mutTypeTable function

The mutTypeTable function summarizes the mutations obtained from mutType
function into a counts table and produces a pdf image plotting the results.
It is automatically run when calling mutType function when parameter
`graphics = T`.

**Usage:** `mutTypeTable(mutTypeResult)`

- `mutTypeResult`: a data.frame object as obtained from the mutType function`

**returns** a table along with a histogram summarizing mutation types

**Example**

```{r mutTypeTable_example}
mutTypeTable(muts)
```

--------------

**Author**:  
Alessandro Giulivo  
Politecnico di Milano  
E-Mail: [alessandro.giulivo@@mail.polimi.it]  
  
Maintainer: Alessandro Giulivo  
  
Github repository: https://github.com/AlessandroGiulivo/sciprog/tree/main/RProj
